Abstractions and exemplars: the measure noun phrase alternation in German

### 2018-01-02 Latex/sections/1intro.tex
Done.

Concerning the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
\footnote{Epistemologically, I believe she refers to acts of \textit{induction} rather than \textit{deduction}.}
It would be highly surprising if this were possible, and the same holds for experimental methods, albeit to a different degree.
Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
Rather, under the standard approach, a theory of cognitive representation is pre-specified.
Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
While the same approach is by and large applicable to corpus data, I now discuss some relevant differences.

@@

I now turn to the second issue, namely which type of cognitive representation research on alternations provides evidence for.
The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of the variants being chosen given these features.
The idea is that a variant is chosen when the influencing features cumulatively assume \textit{prototypical} values for that variant.
In other words, a variant of prototype theory with features \citep{Rosch1978} could be applied, and some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
Prototype theory and exemplar theory model essentially the same types of effects but differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
Similarity-based categorisation is seen as the central and conceptually sufficient approach to linguistic categorisation by many cognitive linguists; see for example \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory.
In the influential framework of cognitive grammar \citep{Langacker1987}, prototypes (which, as should be remembered, represent abstractions already in cognitive science) are literally taken as prototypical exemplars, and there is an additional level of fully discrete abstractions in the form of schemas.
Schemas are characterised by the properties common to all members of the category, whereas a prototypical category member might have very specific additional properties not at all shared by all or even most members \citep[371-375]{Langacker1987}.
The prototype can serve as a reference point when classifying new objects which do not share all properties of the schema, but this would (if repeated) lead to the creation of an even more abstract (hierarchically higher and less specific) schema which describes the new member and the ones belonging to the previous schema.
As pointed out by \citet[136--137]{Langacker1987}, schemas and prototypes thus fulfil different roles and can be assumed to co-exist.
A strict exemplar view of language is incompatible, as far as I can see, with Langacker's view of schemas, but any theory of categorisation that allows for at least some kind of abstraction, is not in principled contradiction with it.%
\footnote{See especially theories with varying intermediate levels of abstraction described below, which seem compatible with Langacker's formation of schema hierarchies through classification by prototype similarity.}
In the remainder of this paper, I do not use schemas in my descriptions of the relevant categories, mostly because the aspect of similarity and fuzzy classification is central to my point, and a formulation in terms of schemas would bring about an unnecessarily high degree of abstraction (see \citealp[70--71]{Taylor2003} for a parallel argument).
However, I fully describe the abstract features of the prototypes, of course, while accounting for item-specific and exemplar effects.

Turning back to the prototype vs.\ exemplar debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones which had been included in the learning data.
This was (at the time) taken as evidence that subjects categorise by prototypes.%
\footnote{See \citet{StormsEa2000} for a comparison of the theories in different experimental settings.}
Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).

In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.%
\footnote{The hard empirical evidence in favour of exemplar models is substantial.
  For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules to be learned.
  This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
For example, \cite{ConawayKurtz2016} show that both prototype theory and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
}
Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.
\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (prototypes) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
\cite{GriffithsEa2009} go further and present a computational model which is able to choose the appropriate complexity of representation for a given category.
However, despite these (and more) attempts to reconcile or unite the two approaches while developing spelled-out mathematical models, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.

>>>>

Concerning the first point, there has for a long time been an interest in correlating probabilistic generalisations extracted from corpus data with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This is often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data'', citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers''.
See also \citet{Newman2011} for a very positive view of corpora as a source of data in cognitive linguistics in their own right.
However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not mean that we can in some way ``deduce mental representations from patterns of use'', \ie\ from corpus data.%
\footnote{Epistemologically, I believe she refers to acts of \textit{induction} rather than \textit{deduction}.}
It would be highly surprising if this were possible, and the same holds for experimental methods, albeit to a different degree.
Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
Rather, under the standard approach, a theory of cognitive representation is pre-specified.
Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted in order to \textit{test} the theory.
While the same approach is by and large applicable to corpus data, I now discuss some relevant differences.

@@

I now turn to the second issue, namely which type of cognitive representation research on alternations provides evidence for.
The typical approach in alternation research is to annotate a large number of corpus sentences with linguistic features and to model the probability of the variants being chosen given these features.
The idea is that a variant is chosen when the influencing features cumulatively assume \textit{prototypical} values for that variant.
In other words, a variant of prototype theory with features \citep{Rosch1978} could be applied, and some researchers such as \cite{Gries2003}, \cite{NessetJanda2010}, or \cite{Schaefer2016c} indeed commit to prototype theory in alternation modelling.
Under prototype theory, category membership is defined by similarity to an ideal exemplar, which is usually identified with characterising features (see \citealp{Taylor2008} for an overview).
While prototype theory is well suited for modeling constructional choices, it is just one of at least two major similarity-based theories of classification, the most prominent other framework being exemplar theory (\citealp{MedinSchaffer1978,Hintzman1986}).
Prototype theory and exemplar theory model essentially the same types of effects but differ significantly in whether they assume higher-level abstractions in the form of single maximally prototypical exemplars or their features (prototype theory) or assume that categories emerge through the storage of many exemplars and similarity classification on those exemplars (exemplar theory).
Similarity-based categorisation is seen as the central and conceptually sufficient approach to linguistic categorisation by many cognitive linguists; see for example \cite{Taylor2003} for a comprehensive treatment in terms of prototype theory.
In the influential framework of cognitive grammar \citep{Langacker1987}, prototypes (which, as should be remembered, represent abstractions already in cognitive science) are literally taken as prototypical exemplars, and there is an additional level of fully discrete abstractions in the form of schemas.
Schemas are characterised by the properties common to all members of the category, whereas a prototypical category member might have very specific additional properties not at all shared by all or even most members \citep[371-375]{Langacker1987}.
The prototype can serve as a reference point when classifying new objects which do not share all properties of the schema, but this would (if repeated) lead to the creation of an even more abstract (hierarchically higher and less specific) schema which describes the new member and the ones belonging to the previous schema.
As pointed out by \citet[136--137]{Langacker1987}, schemas and prototypes thus fulfil different roles and can be assumed to co-exist.
A strict exemplar view of language is incompatible, as far as I can see, with Langacker's view of schemas, but any theory of categorisation that allows for at least some kind of abstraction, is not in principled contradiction with it.%
\footnote{See especially theories with varying intermediate levels of abstraction described below, which seem compatible with Langacker's formation of schema hierarchies through classification by prototype similarity.}
In the remainder of this paper, I do not use schemas in my descriptions of the relevant categories, mostly because the aspect of similarity and fuzzy classification is central to my point, and a formulation in terms of schemas would bring about an unnecessarily high degree of abstraction (see \citealp[70--71]{Taylor2003} for a parallel argument).
However, I fully describe the abstract features of the prototypes.

Turning back to the prototype vs.\ exemplar debate, \cite{Barsalou1990} already showed that prototype and exemplar theory model the same types of effects and are informationally equivalent.
Consequently, experiments which favour one theory over the other use procedural behaviour of subjects in experiments, for example the speed of category retrieval, as opposed to mere output data.
In very early experiments, \cite{PosnerKeele1968} showed, for example, that highly prototypical unseen exemplars were categorised more easily by subjects compared to less prototypical ones which had been included in the learning data.
This was (at the time) taken as evidence that subjects categorise by prototypes.%
\footnote{See \citet{StormsEa2000} for a comparison of the theories in different experimental settings.}
Since corpus data only show artefacts of production events and we have no experimental access to the speaker's or writer's performance and their actual similarity judgements, one should be sceptical whether corpus analysis alone could ever decide which theory of mental representation is more suitable (see also \citealp[22]{Gries2003} and the \citealp[486--487]{Dabrowska2016} quote above).
However, as I will show, some effects are more naturally analysed as prototype effects while others are almost necessarily exemplar effects.

In cognitive science, it is mostly accepted that exemplar theories have greater explanatory power \citep[184]{Vanpaemel2016}, and that abstraction is only needed marginally, if at all.%
\footnote{The hard empirical evidence in favour of exemplar models is substantial.
  For example, in \cite{HahnEa2010}, the authors show that subjects even use exemplar similarity over abstract knowledge even when they are given very simple explicit rules to be learned.
  This is highly relevant because most other studies focus on the learning of implicit rule-based knowledge, which involves many auxiliary assumptions in actual experiments \citep[2]{HahnEa2010}.
On the other hand, there is evidence that neither theory is fully adequate to model humans' capabilities to form categories.
For example, \cite{ConawayKurtz2016} show that both prototype theory and exemplar theory fail to explain certain experimental results where subjects learn to generalise beyond the input in a way that cannot be explained by similarity.
}
Still, various attempts have been made over the past decades to settle the dispute between abstraction-based models (models with rules or prototypes) and exemplar models or to find models which unite the two extremes.
\cite{VanpaemelStorms2008} and \citet{LeeVanpaemel2008} proposed the \textit{varying abstraction model} (VAM) which ``attempt[s] to balance economy and informativeness'' \citep[745]{LeeVanpaemel2008}, treating models with full abstraction (radical prototype theory) and no abstraction at all (radical exemplar theory) as special cases of a model which allows for both abstraction and exemplar effects.
The mixture model of categorisation (MMC) by \cite{Rosseel2002} is a model with abstraction in the form of hierarchical clusters of exemplars, and these clusters of objects are characterised by a probability distribution over their features, and categorising new objects is a process of estimating the probability of this object belonging to one of the clusters.
\cite{GriffithsEa2009} go further and present a computational model which is able to choose the appropriate complexity of representation for a given category.
However, despite these (and more) attempts to reconcile or unite the two approaches while developing spelled-out mathematical models, \cite[183--184]{Vanpaemel2016} describes the state of affairs between adherents of neo-prototype theory (such as \citealp{MindaSmith2001,MindaSmith2002}) and exemplar theory as a stalemate.

### 2018-01-02 Latex/sections/3corpus.tex
Done.

The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
They were calculated from the auxiliary samples described at the end of Section~\ref{sec:gettingdata} as a log-transformed quotient.
The higher the value, the more often the noun occurs in the \PGCd\ (proportionally).
It could be argued that other measures of attraction strength could be used, for example those popularised in collostructional analysis (\citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature), given the overall frequency of the constructions and the lemmas, are cognitively relevant in this case.
After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Sections~\ref{sec:corpushierarchicalmodel} and~\ref{sec:interpretation:corpus}.
Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
To capture this, class information was annotated for measure lemmas as \textit{Measureclass}.
The variable \textit{Cardinal} encodes whether the MNP is modified by a cardinal.

@@

The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).

@@

The main prototypical factors \textit{Cardinal} and \textit{Measureclass} bot pass the PBmodcomp test at sig=0.05 and play the expected role.
The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:conclusion}.

@@

@@

@@

Standard diagnostics show that the model quality is quite good.%
\footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
I am convinced that the model should be specified based on theoretical considerations (and known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
Removal of useless regressors (judging by, for example, the PBmodcomp test and coefficients of determination) is the only type of model selection I allow in my research.}
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and the highest corrected $\text{GVIF}^{1/2\text{df}}$ was 1.520 for \textit{Cardinal}.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
Only \textit{Kindfreq} ($\mpPB=0.095$) could be seen as slightly too high to be convincing, failing at sig=0.05.

Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.%
\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.

@@

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa\, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa.
However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at sig=0.05).
The effect can therefore be ignored or treated as a nuisance variable.

@@

>>>>

The variables \textit{Kindattraction} and \textit{Measureattraction} encode the ratio with which a given kind noun lemma or measure noun lemma occurs in the \PGCd\ and the \NACb.
They were calculated from auxiliary samples to be described in Section~\ref{sec:mainstudysampling} as a log-transformed quotient.
The higher the value, the more often the noun occurs in the \PGCd\ (proportionally).
It could be argued that other measures of attraction strength could be used, for example those popularised in collostructional analysis (\citealp{StefanowitschGries2003,GriesStefanowitsch2004}; see also \citealp{Gries2015a}).
However, the goal here is to quantify how often lemmas occur in the \PGCd\ and the \NACb, and these constructions do not compete at all but are rather mutually exclusive.
While this does not preclude the use of collostructional analysis, it is an open question whether the marginals (usually called the ``expected frequencies'' in the collostructional literature), given the overall frequency of the constructions and the lemmas, are cognitively relevant in this case.
After all, the main difference between the quotient used here and the logarithmised p-values from a Fisher test is that they take these marginals into account. 
However, using collexeme strength instead of the raw frequency quotient was tried as an alternative (variables \textit{Kindcollo} and \textit{Measurecollo}); see Section~\ref{sec:corpushierarchicalmodel} for a discussion of the negative results.

Additionally, \textit{Kindfreq} and \textit{Measurefreq} are the logarithm-transformed frequencies per 1,000,000 words of each lemma, extracted from the frequency lists distributed by the DECOW14A corpus creators on their web page.
In Section~\ref{sec:analyses}, it was hypothesised that classes of measure lemmas might have different preferences for the two alternants.
To capture this, class information was annotated for measure lemmas as \textit{Measureclass}.
The variable \textit{Cardinal} encodes whether the MNP is modified by a cardinal.

@@

The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and consequently have no coefficient estimates.%
\footnote{My approach to statistical inference is essentially Fisherian \citep{Fisher1935,Fisher1959}, and hence I use the term \textit{sig level} instead of the term \textit{$\alpha$ level} known from null hypothesis significance testing (NHST).
For overviews of systems of statistical inference with special focus on this and related questions, see \citet{Lehmann1993,Lehmann2011} or the shorter \citet{Perezgonzalez2015}.}
The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).

@@

The main prototypical factors \textit{Cardinal} and \textit{Measureclass} bot pass the PBmodcomp test at sig=0.05 and play the expected role.
The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:mainstudyinterpretation}.

@@

\subsubsection{Sampling and annotation}
\label{sec:mainstudysampling}

@@

\subsubsection{Statistical model}
\label{sec:mainstudystatisticalmodel}

@@

Standard diagnostics show that the model quality is quite good.%
\footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
I am convinced that models should be specified based on theoretical considerations (including known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
Removal of useless regressors (judging by, for example, the PBmodcomp test and coefficients of determination) is the only type of model selection I allow in my research.}
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and the highest corrected $\text{GVIF}^{1/2\text{df}}$ was 1.520 for \textit{Cardinal}.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
Only \textit{Kindfreq} ($\mpPB=0.095$) could be seen as slightly too high to be convincing, failing at sig=0.05.

Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid \textit{log(0)} errors) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
The coefficients of determination drop to $R_m^2=0.376$ and $R^2_c=0.480$.

\subsubsection{Interpretation}
\label{sec:mainstudyinterpretation}

@@

An interesting picture emerges for the lemma frequencies.
A higher-than-average lemma frequency of measure nouns favours the \NACa, which is as expected if we assume at least a tendency for highly grammaticalised items to be more frequent.
With kind nouns, higher frequency seems to favour the \PGCa.
However, there is no clear theoretical interpretation (see Section~\ref{sec:analyses}), and the estimate is imprecise (not significant at sig=0.05).
The effect can therefore be ignored or treated as a nuisance variable.

@@

To close this section, I now compare the results of the pre-study and the main study.
Even though the non-alternating constructions are surely subject to additional constraints (especially because they are much less interchangeable than the alternating constructions), the coefficients align neatly in many cases.
First of all, the intercepts encode a similar overall dominance of the NAC constructions (pre-study: -5.370; main study: -3.548).
For the \textit{Cardinal} effect, the coefficients 1.419 (pre-study) and 1.189 (main study) have the same sign and magnitude and mostly overlapping confidence intervals.
The levels of \textit{Measureclass} have comparable coefficients, although the divergence is larger.
In both studies, the abstract measure nouns in the \textit{Physical} class (on the intercept) are most clearly associated with the NAC constructions.
Also, in both cases, the \textit{Container} class is closest to \textit{Physical} with the same sign and magnitude (pre-study: 0.445; main study: 0.252).
The main difference -- if we ignore the \textit{Rest} class which can be expected to show no clear tendency -- is that \textit{Amount} and \textit{Portion} are slightly closer together in their tendency to favour the PGC constructions in the pre-study (\textit{Amount} 1.597 and \textit{Portion} 1.782 in the pre-study vs.\ \textit{Amount} 0.831 and \textit{Portion} 1.217 in the main study).
The difference is not huge, and the overall order of the coefficients is the same.
The \textit{Genitives} effect also converges with -0.710 in the pre-study and -0.693 in the main study.
The two studies do not converge with respect to the \textit{Badness} variable, which is not significant in the pre-study.
It would, however, surprise to achieve perfectly converging results given that even real effects are missed at certain rates.
In the case of \textit{Badness}, we see that even in the main study, the effect size is small (-0.152), and thus even at the impressive sample size of roughly 5,000 in both studies, the effect might simply be too weak to be detected reliably.
Finally, the frequency effect \textit{Measurefreq} is detected in the main study but not in the pre-study, while the \textit{Kindfreq} effect is essentially absent in both studies.
In connection with this, it is revealing to look at the coefficients of determination.
In the pre-study, a much greater proportion of the variance is explained by taking the lemma random effects into account ($R^2_m=0.278$ and $R^2_c=0.566$) than it is in the main study ($R^2_m=0.376$ and $R^2_c=0.480$).
Thus, the frequency effect for measure lemmas might be swamped by the lemma random intercepts.
All things considered, the studies have shown that the predicted prototype effects are mostly found in usage data for both the alternating and the non-alternating constructions.
